
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Extra Credits"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment4_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- Notice that if there is `eval=FALSE` in the first line of the code chunk, the chunk will not be execute.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

This assignment works with the [IMDB Top 1000 data](../data/imdb_top_1000.csv).  Find out more information about this data at [this link](https://www.kaggle.com/datasets/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows).  Import the data and answer the following questions. 
```{r}
library(tidyverse)
df <- read_csv("imdb_top_1000.csv")
```

1. List all the names of the columns of the data
```{r}
names(df)
```

2. Which movies have the highest money earned (Gross)?
```{r}
# Producing a table of movie titles and their money earned (gross) ordered...
#...from highest to lowest gross
df %>% select(Series_Title, Gross) %>% arrange(-Gross)
```

3. What is the lowest rating (IMDB_Rating)?  List five movies have this lowest rating.
```{r}
# Getting lowest IMDB_Rating in the data
min(df$IMDB_Rating)

# Producing list of movie titles and IMDB_Ratings where IMDB_Rating is equal...
#...to 7.6 (lowest rating in data found above)
df %>% select(Series_Title, IMDB_Rating) %>% filter(IMDB_Rating == 7.6)

# 5 movies with a low rating of 7.6 include:
  # Dark Waters
  # Searching
  # Once Upon a Time... in Hollywood
  # Nelyubov
  # The Florida Project
```

4. Which year have the most number of movies released in the list?  What is the total of money earned on that year?
```{r}
# Getting list of release years and number of movies produced during...
#...those years
df %>% count(Released_Year) %>% arrange(-n)

# 2014 had the most number of movies released in the list with 14

# Getting total money earned during 2014
df %>% filter(Released_Year == 2014) %>% summarise(Total_Money = sum(Gross, na.rm=TRUE))

# In total, $2,755,629,221 was earned in 2014
```

5. What is the total money earned per movies on average?
```{r}
df %>% summarise(Average_Total_Money = mean(Gross, na.rm=TRUE))

# On average, the total money earned per movie is $68,034,751
```

6. Calculate the average number of votes by year. Calculate the average number of votes of movies that have IMDB rating greater than 9.
```{r}
# Calculating average number of votes by year
df %>% group_by(Released_Year) %>% summarise(Average_Votes = mean(No_of_Votes))

# Calculating average number of votes by year for movies w/ IMDB rating > 9
df %>% filter(IMDB_Rating > 9) %>% 
  group_by(Released_Year) %>% 
  summarise(Average_Votes = mean(No_of_Votes))
```

7. Calculate the average Meta score in 2020 of movies that have number of votes in the third quartile. 
```{r}
# Assuming "in the third quartile" means vote totals between total that is > 50% of ... 
# ...the data and the total that is >75% of the data

df %>% filter(No_of_Votes > quantile(df$No_of_Votes, probs=0.50) | 
                No_of_Votes < quantile(df$No_of_Votes, probs=0.75),
              Released_Year == 2020) %>%
  summarise(Avg_Meta_Score = mean(Meta_score, na.rm=TRUE))
```

8. (Optional - Challenging). The current `Runtime` variable is not a numeric. Use the `str_remove` function to remove ` min` from the variables then use `as.numeric` to convert the variable to numeric. Calculate the average running time in the 2010s. Calculate the correlation between running time and rating (adding `use="complete.obs"` in the `cor` function to ignore the missing values).  
```{r}
# Removing 'min' from the runtime observations
df$Runtime <- str_remove(df$Runtime, 'min')

# Converting runtime observations to numeric
df$Runtime <- as.numeric(df$Runtime)

# Calculating average running time in the 2010s
df %>% filter(Released_Year >= 2010 | Released_Year < 2020) %>%
  summarise(Avg_Runtime = mean(Runtime))

# Calculating correlation between running time and rating
cor(df$Runtime, y = df$IMDB_Rating, use='complete.obs')
```

9. We can use `select_if` to select columns satisfying a condition and use `summarise_if` to do calculation on columns satisfying a condition.  Try the follows to understand these functions. 

```{r,eval=FALSE}
# Select only character columns
df %>% select_if(is.character)

# Calculate the median of all numeric columns
df %>% summarise_if(is.numeric, mean, na.rm=TRUE)
```


10. Implement the follows functions or combos. Drawing a comment or summary from each calculation. The codes in this question should be different from the codes used in other questions.

- select
- filter
- mutate
- summarise
- arrange
- count
- count + arrange
- filter + count + arrange
- group_by + summarise
- filter + group_by + summarise
- filter + group_by + summarise + arrange
```{r}
# Selecting movie titles and their runtimes
df %>% select(Series_Title, Runtime)
```
```{r}
# Filtering df to only include movies released in 2005
df %>% filter(Released_Year == 2005)
```
```{r}
# Creating variable to see the money earned per minute of each film
df %>% mutate(gross_per_minute = Gross / Runtime)
```
```{r}
# Calculating average runtime of all movies in the data
df %>% summarise(Avg_Runtime = mean(Runtime))
```
```{r}
# Arranging rows in df by number of votes in descending order
df %>% arrange(-No_of_Votes)
```
```{r}
# Generating number of movies by genre
df %>% count(Genre)
```
```{r}
# Counting number of movies for each certificate type and arranging them in...
# ...descending order
df %>% count(Certificate) %>%
  arrange(-n)
```
```{r}
# Determining which director has the most movies that run longer than 90 min...
# ...in the dataset
df %>% filter(Runtime > 90) %>%
  count(Director) %>%
  arrange(-n)
```
```{r}
# Determining the average IMDB rating for each Star1 (lead) actor/actress...
# ... and the films they starred in
df %>% group_by(Star1) %>%
  summarise(Avg_IMDB_Rating = mean(IMDB_Rating))
```
```{r}
# Determining median meta score for the performances Star2 actors/actresses...
# ...in Drama films
df %>% filter(Genre == 'Drama') %>%
  group_by(Star2) %>%
  summarise(Median_Meta_Score = median(Meta_score))
```
```{r}
# Determining the average runtime of films directed by Steven Spielberg based on...
# ...genre and ordering the rows in descending order
df %>% filter(Director == 'Steven Spielberg') %>%
  group_by(Genre) %>%
  summarise(Avg_Runtime = mean(Runtime)) %>%
  arrange(-Avg_Runtime)
```

