
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7: Visualization - Some aspects of ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

-------

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis. 

-------

1. Make a plot using `facet_wrap` by one categorical variable (`facet_wrap(~variable1)`)
```{r}
# Calling libraries
library(tidyverse)
library(dplyr)
library(ggplot2)

# Reading in data
df <- read_csv('spotify-2023.csv')

# Creating plot to compare the number of times a song appeared in Apple playlists and Spotify playlists across song mode
ggplot(df, aes(x=in_spotify_playlists, y=in_apple_playlists)) + geom_point(color='Green') + facet_wrap(~mode, ncol=3) + theme_dark() + ggtitle("Comparison of Number of Times a Song Appeared in Apple vs Spotify Playlists\nAcross Song Mode (Major vs Minor)") + labs(caption = "Based on 2023 Spotify Data") + theme(plot.caption = element_text(hjust=1)) 

# Regardless of song mode, the correlation between the number of times a song appears on Apple and Spotify playlists appears similar (both relatively positive). However this correlation appears to be slightly stronger for Major songs.

# The commented out line of code below was used to save this plot so I could place it on my Github site
# ggsave("my_saved_plot.png")
```

2. Make a plot using `facet_wrap` by two categorical variables (`facet_wrap(~variable1+variable2)`)
```{r}
# Creating a filtered df to only include songs that were written in the keys A or A#
filtered_df_1 <- df %>% filter(substr(df$key,1,1) == 'A')

# Creating a plot to compare the positioning of a song in Apple vs Spotify charts based on the song mode and key
ggplot(filtered_df_1, aes(x=in_spotify_charts, y=in_apple_charts)) + geom_point(color='Green') + facet_wrap(~key+mode) + theme_dark() + ggtitle("Comparison of Positioning of a Song in Apple vs Spotify Charts\nby Song Mode (Major vs Minor) and Key (A vs A#)") + labs(caption = "Based on 2023 Spotify Data") + theme(plot.caption = element_text(hjust=1))

# It appears that the strongest positive correlation between song positioning in Apple and Spotify charts occurs with minor songs that were written in the A key. The weakest correlation appears to occur with minor songs written in the A# key.
```

3. Make a plot using `position ='dodge'`.
```{r}
# Filtering df to only incldue songs written in keys D & D#
filtered_df_2 <- df %>% filter(substr(df$key,1,1) == 'D')

# Plotting the frequency of songs based on their key (D &) and mode
ggplot(filtered_df_2, aes(x=key, fill=mode)) + geom_bar(position = 'dodge', na.rm = TRUE, color='Black') + theme_dark() + ggtitle("Frequency of Spotify Songs by Key (D vs D#) and Mode (Major vs Minor)") + labs(caption = "Based on 2023 Spotify Data") + theme(plot.caption = element_text(hjust=1))

# The majority of songs in this subset of the data are major songs written in the D key. Interestingly, even though there are many more major songs written in the D key, there are actually more minor songs than major written in the D# key.
```

4. Make a plot using `position ='fill'`. 
```{r}
# Filtering df to only include songs written in F and F# keys
filtered_df_3 <- df %>% filter(substr(df$key,1,1) == 'F')

# Creating a plot to show the proportion of songs by mode and key (F & F#)
ggplot(filtered_df_3, aes(x=mode, fill=key)) + geom_bar(position = 'fill', na.rm = TRUE, color='Black') + theme_dark() + ggtitle("Proportion of Spotify Songs by Mode (Major & Minor) and Key (F & F#)") + labs(caption = "Based on 2023 Spotify Data") + theme(plot.caption = element_text(hjust=1)) + labs(y='Proportion')

# It appears that the proportion of minor songs is nearly evenly split between being written in the F & F# keys, while there is a less even split for major songs (with around 60% of these songs being written in the F key and only around 40% written in F#).
```

5. Make a plot using `geom_col`.
```{r}
# Filtering df to only include songs written in the G & G# keys
filtered_df_4 <- df %>% filter(substr(df$key,1,1) == 'G')

# Plotting the number of times a song appeared in Deezer playlists based on their mode and key (G or G#)
ggplot(filtered_df_4, aes(x=mode, y=in_deezer_playlists, fill=key)) + geom_col(na.rm = TRUE, position = 'dodge') + theme_light() + ggtitle("Number of Times Spotify Songs Appear in Deezer Playlists by Mode (Major & Minor)\nand Key (G & G#)") + labs(caption = "Based on 2023 Spotify Data") + theme(plot.caption = element_text(hjust=1))

# Many more major songs written in the G and G# keys appeared in Deezer playlists in comparison to the minor songs written in G and G#. Interestingly as well, songs written in the G key were in more Deezer playlists for major songs, while more minor songs written in the G# key were found in Deezer playlists.
```

6. Change the theme of one of your plots. 
```{r}
# Changed the theme of my plot for #3 to theme_void()
ggplot(filtered_df_2, aes(x=key, fill=mode)) + geom_bar(position = 'dodge', na.rm = TRUE, color='Black') + theme_void() + ggtitle("Frequency of Spotify Songs by Key (D vs D#) and Mode (Major vs Minor)") + labs(caption = "Based on 2023 Spotify Data") + theme(plot.caption = element_text(hjust=1))

# This theme does not seem overly useful as it strips away the axes, intervals, and grid lines :)
```

7. Make a plot using combo filter + ggplot
```{r}
# Comparing the number of song appearances in Spotify vs Apple playlists for The Weeknd
df %>% filter(df['artist(s)_name'] == 'The Weeknd') %>% ggplot() + geom_point(mapping = aes(x=in_apple_playlists, y=in_spotify_playlists), color = 'Green') + ggtitle("Comparison of Song Appearances in Spotify vs Apple Playlists for The Weeknd") + theme_dark() + labs(caption = "Based on 2023 Spotify Data") + theme(plot.caption = element_text(hjust=1))

# The Weeknd's songs appear to follow the general trend of all songs when comparing appearances in Spotify and Apple playlists, as it as they display a positve correlation (the more times a song appears in Apple playlists, the more times they appear in Spotify playlists). 
```

8. Make a plot using combo group_by + summarise + geom_col
```{r}
# Creating a chart to look at the average beats per minute of a song based on the key in which the song was written
df %>% filter(!is.na(key)) %>% group_by(key) %>% summarise(avg_bpm = mean(bpm)) %>% ggplot() + geom_col(mapping = aes(x = key, y = avg_bpm), fill = 'Green', color = 'Black') + theme_dark() + ggtitle("Average Beats per Minute of Spotify Songs Based on their Key") + labs(caption = "Based on 2023 Spotify Data") + theme(plot.caption = element_text(hjust=1))

# Regardless of song key, the average beats per minute of a song appears to remain similar (around 120-125 bpm). Songs written in the A key appear to have the highest average bpm, with around 130 bpm.
```

9. Save a plot and put a link of the plot to your Github Webpage.  To link a plot `abc.png` you just need to insert the following on the README.md

```{r, eval=FALSE}
# Decided to save my plot from #1 for this problem (link is visible on main Github site underneath the link for this assignment)
[Link to the plot](abc.png)
```

10. In this question, we will use  the Covid19 vaccination data for the US. The data and data dictionary can be found at the below link:

- [Data](https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv)

- [Alternative Link](https://bryantstats.github.io/math421/data/vaccinations-by-manufacturer.csv)

```{r}
library(tidyverse)
df <- read_csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv')
```



`Geom_point` makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types used in the US by

```{r}
df %>% 
  filter(location=='United States') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))
```

Do the follows. 

  - Use `geom_point` to compare the vaccines used by date in Argentina in 2022. 
```{r}
# Creating year column
library(lubridate)
df$year<- year(df$date)

# Creating a plot to compare the vaccines used by date in Argentina in 2022
df %>% 
  filter(location=='Argentina', year == 2022) %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine)) + labs(caption = "Based on 2020-2022 Covid-19 Vaccination Data") + theme(plot.caption = element_text(hjust=1))

# It appears that for the entirety of 2022, the vaccine from Sinopharm/Beijing was the most common vaccine being used in Argentina, while the least utilized was the CanSino vaccine. Throughout 2022, the Moderna vaccine saw the biggest uptake as at the start of the year there were around 4 million Moderna vaccines used, while by the end of the year there were around 16 million used.
```
  
  - Change `geom_point` to `geom_line` in the above codes to have the corresponding line plots. 
```{r}
# Creating a line plot to show the vaccine utilization trends in Argentina in 2022
df %>% 
  filter(location=='Argentina', year == 2022) %>% 
  ggplot()+
  geom_line(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine)) + labs(caption = "Based on 2020-2022 Covid-19 Vaccination Data") + theme(plot.caption = element_text(hjust=1))

# The comments from the scatter plot above also apply to this plot as they visualize the same trend, with the only difference being that this plot is a line plot while the plot above was a scatter plot.
```

11. (Using the same dataset as in 10.) Sometime we need to do some calculations on the data before plotting. For example, the below codes is to compare the total vaccines (including all types of vaccines) by date between the US and the EU

```{r}
# the total vaccines (including all types of vaccines) by date
d1 <- df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE))

# Plot
d1 %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```

Notice that the above codes could be piped in one 

```{r, eval=FALSE}
df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```


Do the follows. 

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in Asia
```{r}
# Creating plot to compare the total vaccines (including all types of vaccines) by date between the countries/regions in Asia [specifically, Cyprus, Hong Kong, Japan, Nepal, and South Korea]
df %>% 
  filter(location %in% c('Cyprus','Hong Kong','Japan','Nepal','South Korea')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location)) + labs(caption = "Based on 2020-2022 Covid-19 Vaccination Data") + theme(plot.caption = element_text(hjust=1))

# Japan had the most total vaccines from around April or May 2021 on, also seeing the most rapid increases in vaccines over this time period. It appears that there was potentially inconsistent reporting from South Korea starting from around February 2022 as the total vaccines in South Korea appears to fluctuate between 90 million, 70 million, and just over 50 million from February 2022 on. Similar issues with reporting can be seen from Nepal as there appears to be a gap in reporting from around March 2021 to February 2022. Although Hong Kong and Cyprus had consistent reporting, their total number of vaccines appear to be much lower over this timeframe.
```

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in South America in 2021
```{r}
# Creating a plot to compare the total vaccines by date between South American countries in 2021 (specifically Argentina, Chile, Ecuador, Peru, and Uruguay)
df %>% 
  filter(location %in% c('Argentina','Chile','Ecuador','Peru','Uruguay'), year == 2021) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location)) + labs(caption = "Based on 2020-2022 Covid-19 Vaccination Data") + theme(plot.caption = element_text(hjust=1))

# By the end of 2022, Argentina had the most total vaccines (around 28 million more than Peru [the next closest country]). This was not always the case however, as halfway through Q1 of 2022, the total number of vaccines in Chile began to increase (meaning that Chile had the most total vaccines out of these countries at that time). Ecuador did not see an increase similar to Chile's until early Q3 in 2022, while Uruguay never saw a dramatic increase in total vaccines during 2022.
```

