
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

**Notice:** *In this assignment, all the plot should have title, caption, and axis labels. *

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths
```{r}
# Calling libraries
library(gganimate)
library(ggplot2)
library(tidyverse)
library(lubridate)

# Reading in Covid-19 WHO data
df <- read_csv('WHO-COVID-19-global-data.csv')

df %>% 
  filter(Date_reported == max(Date_reported)) %>%
  arrange(-Cumulative_deaths) %>%
  filter(rank(-Cumulative_deaths) <= 3)

# The three countries with the most number of deaths by Covid-19 are the United States of America, Brazil, and India.
```


2. Use `transition_reveal` to make a plot revealing the number of deaths in the three countries with the most numbers of deaths
```{r}
# Creating plot revealing the number of deaths in the 3 countries with the most numbers of deaths
plot_1 <- df %>% 
  filter(Country %in% c('United States of America', 'Brazil', 'India')) %>%
  ggplot(aes(x=Date_reported, y=Cumulative_deaths, color=Country)) + geom_line() + transition_reveal(Date_reported) + ggtitle("Number of Covid-19 Deaths in the USA, Brazil,\nand India from 2020-2023") + labs(caption = "Based on World Health Organization (WHO)\ndata collected from 2020-2023") + theme(plot.caption = element_text(hjust=1)) + theme(plot.title = element_text(hjust = 0.5)) 

animate(plot_1)
```

3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 
```{r}
df$death_per_cases <- (df$Cumulative_deaths/df$Cumulative_cases)

df %>% 
  filter(Date_reported == max(Date_reported)) %>%
  arrange(-death_per_cases) %>%
  filter(rank(-death_per_cases) <= 3)

# Yemen, Sudan, and the Syrian Arab Republic are the three countries with the highest deaths per cases.
```

4. Use `transition_reveal` to make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 
```{r}
# Creating plot to reveal the number of deaths per cases of the US, Italy, and Mexico
plot_2 <- df %>% 
  filter(Country %in% c('United States of America', 'Mexico', 'Italy')) %>%
  ggplot(aes(x=Date_reported, y=death_per_cases, color=Country)) + geom_line(na.rm = TRUE) + transition_reveal(Date_reported) + ggtitle("Number of Covid-19 Deaths per Cases\nin the USA, Mexico, and Italy from 2020-2023") + labs(caption = "Based on World Health Organization (WHO)\ndata collected from 2020-2023") + theme(plot.caption = element_text(hjust=1)) + theme(plot.title = element_text(hjust = 0.5)) 

animate(plot_2)
```

5. Import the following data https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv
                      
Use `transition_reveal` to plot the total vaccinations per hundred by level of income. Hint: the levels of income are identified in the `location` variable. Put a short comment on the caption of the plot. 
```{r}
# Reading in the Covid-19 Vaccination data
df2 <- read_csv('vaccinations.csv')

# Creating plot to show the total vaccinations per hundred by income level
plot_3 <- df2 %>% 
  filter(location %in% c('High income', 'Low income', 'Lower middle income', 'Upper middle income')) %>%
  mutate(date = as.Date(as.character(mdy(date)))) %>%
  ggplot(aes(x=date, y=total_vaccinations_per_hundred, color=location)) + geom_line() + transition_reveal(date) + ggtitle("Number of Total Vaccinations (per Hundred) by Level\nof Income from 2021-2023") + labs(caption = "Based on Our World in Data (OWID)\ndata collected from 2021-2023\n\nThis plot indicates that there appears to be a correlation between income\nlevels and total vaccinations (per Hundred), as by September 2023,\nthe higher the income level, the more total vaccinations (per hundred).") + theme(plot.caption = element_text(hjust=1)) + theme(plot.title = element_text(hjust = 0.5))

animate(plot_3)
```

