---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*****:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***:

-   Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio.

-   Right under each question, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3.

------------------------------------------------------------------------

## Problems

\

1.  Create the following data frame

| Rank | Age | Name  |
|:-----|:----|:------|
| 0    | 28  | Tom   |
| 1    | 34  | Jack  |
| 2    | 29  | Steve |
| 3    | 42  | Ricky |

```{r}
df <- data.frame(Rank = c(0, 1, 2, 3),
                 Age = c(28, 34, 29, 42), 
                 Name = c("Tom", "Jack", "Steve", "Ricky"))
```

2.  Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).

```{r}
data <- read.csv("us-states.csv")
```

2.  Show the names of the variables in the data

```{r}
names(data)
```

3.  How many columns and rows the data have?

```{r}
ncol(data)
nrow(data)
```

4.  How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?
```{r}
sum(is.na(data))
```

5.  What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column.
```{r}
class(data$date)
```

6.  Capitalize the names of all the variables
```{r}
names(data) <- toupper(names(data))
```

7.  Find the average number of cases per day. Find the maximum cases a day.
```{r}
mean(data$CASES)
max(data$CASES)
```

8.  How many states are there in the data?
```{r}
length(unique(data$STATE))
```

9.  Create a new variable `weekdays` to store the weekday for each rows.
```{r}
data$DATE = as.Date(data$DATE)
data$weekdays <- weekdays(data$DATE)

# Creating a month column in this block to make #15 easier
#install.packages("lubridate")
library(lubridate)

data$month <- month(ymd(data$DATE))
```

10. Create the categorical variable `death2` variable taking the values as follows

-   `has_death` if there is a death that day
-   `no_death` if there is no death that day

Find the frequency and relative frequency of `no_death` and `has_death`.
```{r}
data$death2 <- ifelse(data$DEATHS == 0, 'no_death', 'has_death')
```

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint: Use the `summary` function)
```{r}
summary(data$DEATHS)
```

12. Create the categorical variable `death3` variable taking the values as follows

-   `low_death` if the number of deaths smaller than the 25 percentile (Q1)

-   `mid_death` if the number of deaths from Q1 to Q3

-   `high_death` if the number of deaths greater than Q3
```{r}
#install.packages("dplyr")
library(dplyr)
data$death3 <- case_when(data$DEATHS < quantile(data$DEATHS, 0.25) ~ 'low_death',
                   data$DEATHS <= quantile(data$DEATHS, 0.75) ~ 'mid-death',
                   TRUE ~ 'high_death')
```

13. Find the average cases in Rhode Island in 2021
```{r}
df_RI_Cases_2021 <- data[data$DATE >='2021-01-01' & data$DATE <= '2021-12-31' & data$STATE == 'Rhode Island',]

mean(df_RI_Cases_2021$CASES)
```

14. Find the median cases by weekdays in Rhode Island in 2021
```{r}
# Finding the median cases in RI during 2021 by weekday
by(df_RI_Cases_2021$CASES,df_RI_Cases_2021$weekdays, median)
```
```{r}
class(data$month)
```

15. Compare the median cases in Rhode Island in June, July, August and September in 2021.
```{r}
# Creating a subset of df_RI_Cases_2021 to only include rows for the...
# ...given months above
df_RI_Summer_Cases_2021 <- df_RI_Cases_2021[df_RI_Cases_2021$month %in% c(6,7,8,9),]

# Finding the median cases in RI during 2021 by month (for the given...
# ...months above)
by(df_RI_Summer_Cases_2021$CASES,df_RI_Summer_Cases_2021$month, median)

```

16. Find your own dataset, import it and implement the following functions on the data

-   head
-   str
-   names
-   mean, min, max
-   table
-   is.na
-   colSums
-   class
-   cor
-   by
-   ifelse
-   case_when
```{r}
titanic_df <- read.csv("titanic.csv")
head(titanic_df)
```
```{r}
str(titanic_df)
```
```{r}
names(titanic_df)
```
```{r}
# Determining what proportion of passengers survived (from the dataset)
mean(titanic_df$Survived)
# Determining the minimum ticket fare
min(titanic_df$Fare)
# Determining the number of passengers (in the dataset)
max(titanic_df$PassengerId)
```
```{r}
class(titanic_df$Age)
```
```{r}
# Counting the number of missing values in the data
sum(is.na(titanic_df))
```
```{r}
# Determining the number of missing values in each column
colSums(is.na(titanic_df))
```
```{r}
# Seeing relationship between ticket fare and passenger class
cor(titanic_df$Fare, titanic_df$Pclass)
```
```{r}
# Determining the survival rate of passengers by their passenger class
by(titanic_df$Survived, titanic_df$Pclass, mean)
```
```{r}
# Creating categorical variable for passenger survival
titanic_df$survived_label <- ifelse(titanic_df$Survived == 0, "died", "survived")
```
```{r}
# Creating categorical variable from numeric variable for Pclass
titanic_df$Pclass_label <- case_when(titanic_df$Pclass == 1 ~ "first class", titanic_df$Pclass == 2 ~ "second class", titanic_df$Pclass == 3 ~ "third class")
```

17. In the dataset in #16, practice the follows. You can reuse the code of 16.

-   Create a categorical variable from a continuous variable
-   From one categorical variable, create a new categorical variable with fewer categories
```{r}
# Creating categorical variable to define ticket fare as cheap or expensive
titanic_df$ticket_price_range <- ifelse(titanic_df$Fare > mean(titanic_df$Fare), "expensive", "cheap")

# Creating a new categorical variable that takes Pclass label variable...
# ...from above cell and reduces categories to "first class" or "economy"
titanic_df$first_class_label <- (ifelse(titanic_df$Pclass_label == "first class", "first class", "economy"))
```

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
