
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 15: Writing a function - Conditional Statement"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment15.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

#### NOTICE: Question 7, 8 and 9 are optional. 

-------

1. Write a function to calculate the age of a person.  Give an example to test your function. 

- Input: a number, x, (year born)

- Output: Print out "You are r age". Where r is the age of the person, i.e. 2020 - x. 

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#42)
```{r}
age <- function(x)
{
  print(paste0("You are ", as.integer(format(Sys.Date(), "%Y")) - x, " years old."))
  }

# Test case: A working function should return "You are 13 years old." as of the creation of this code (2023)
age(2010)

```

-------

2. Write the following function. Give an example to test your function. 

- Input: a number

- Output: print out: "You input an even number!" if the number is even, or "You input an odd number!" otherwise. 

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#46)
```{r}
even_or_odd <- function(x)
{
  if (x %% 2 == 0)
{print("You input an even number!")}
  else
 {print("You input an odd number!")}
}

# Test case: A working function should return "You input an odd number!"
even_or_odd(3)

# Test case: A working function should return "You input an even number!"
even_or_odd(96)
```

-------

3. Write the following function.  Give an example to test your function. 

- Input: a numeric vector 

- Output:  

  - if the input vector has missing values: return the input vector with missing values replaced by mean
  
  - if the input vector has no missing value: return the same input vector
  
Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#50)
```{r}
mean_impute <- function(x)
{
mean_of_x <- mean(x, na.rm = TRUE)
library(tidyr)
x <- replace_na(x, mean_of_x)
return(x)
}

# Test case: A working function should return a numeric vector containing c(1, 2, 3, 2)
mean_impute(c(1, NA, 3, NA))

# Test case: A working function should return a numeric vector containing c(1, 2, 3, 4)
mean_impute(c(1, 2, 3, 4))
```

-------

4. Write the following function.  Give an example to test your function. 

- Input: a vector x 

- Output:  The vector x where the missing values replaced by the mean (if x is numeric) or the mode (if x is non-numeric).  If x does not have missing value, return the same vector x.  

Hint: Use If-statement to combine the function in Question 3 and [this function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#49)
```{r}
vector_impute <- function(x)
{
  if(!is.numeric(x))
  {
    # Find the mode of x
    mode_of_x <- names(sort(-table(x)))[1]
    # Replace the missing by the mode
    library(tidyr)
    x <- replace_na(x, mode_of_x) 
  }
  else
  {
    mean_of_x <- mean(x, na.rm = TRUE)
library(tidyr)
x <- replace_na(x, mean_of_x)
}
return(x)    
}

# Test case: A working function should return a non-numeric vector of c("a", "a", "b", "a")
vector_impute(c("a", "a", "b", NA))

# Test case: A working function should return a numeric vector of c(1, 4, 7, 4)
vector_impute(c(1, NA, 7, NA))

# Test case: A working function should return a non-numeric vector of c("a", "b", "c", "d")
vector_impute(c("a", "b", "c", "d"))

# Test case: A working function should return a numeric vector of c(2, 4, 6, 8)
vector_impute(c(2, 4, 6, 8))
```

-------

5. Write the following function. Give examples to test your function. 

- Input: A data frame of two variables x and y

- Output: 

  - A boxplot of x by y if x is numeric and y is non-numeric
 
  - A boxplot of y by x if y is numeric and x is non-numeric
 
  - print out 'This function cannot visualize your data' otherwise

Hint: 

- [Similar Function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#53)
 
- You can refer to this slide to plot a boxplot: https://bryantstats.github.io/math421/slides/6_viz.html#36
```{r}
create_boxplot <- function(d)
{ 
  library(ggplot2)
  if (is.numeric(d[[1]]) && !is.numeric(d[[2]]))
  {
    d %>% ggplot() + geom_boxplot(mapping = aes(x = d[[1]], y = d[[2]])) + labs(y=colnames(d[2]), x =colnames(d[1]))
  }
  else if (is.numeric(d[[2]]) && !is.numeric(d[[1]]))
  {
    d %>% ggplot() + geom_boxplot(mapping = aes(x = d[[2]], y = d[[1]])) + labs(y=colnames(d[1]), x =colnames(d[2]))
  }
  else
  {
    print("This function cannot visualize your data")
  }
}

library(tidyverse)
df <- read_csv("titanic.csv")
df = drop_na(df)

# Test case: A working function should return a box plot with Age along the x-axis and the categories for Embarked along the y-axis
create_boxplot(df %>% select(Age, Embarked))

# Test case: A working function should return a box plot with Age along the x-axis and the categories for Age along the y-axis (this example shows that the order of the variables does not matter, so long as one is numeric and one is not numeric)
create_boxplot(df %>% select(Sex, Age))

# Test case: A working function should return 'This function cannot visualize your data'
create_boxplot(df %>% select(Name, Sex))
```
 
-------
 
6. Combine the function in Question 5 and the function in this example ([Link](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#53)) to have a function that can plot all possible cases (2 categorical, 2 numeric, 1 numeric and 1 categorical). 
```{r}
visualize <- function(d)
{tryCatch(
  {if(is.numeric(d[[1]])&is.numeric(d[[2]]))
    
  {
    d %>% ggplot(aes(x = d[[1]], y = d[[2]]))+
      geom_point()+
      labs(x = names(d)[1], y = names(d)[2])
  }
  
  else if (is.numeric(d[[1]]) && !is.numeric(d[[2]]))
    
  {
    d %>% ggplot() + geom_boxplot(mapping = aes(x = d[[1]], y = d[[2]])) + labs(y=colnames(d[2]), x =colnames(d[1]))
  }
  
  else if (is.numeric(d[[2]]) && !is.numeric(d[[1]]))
    
  {
    d %>% ggplot() + geom_boxplot(mapping = aes(x = d[[2]], y = d[[1]])) + labs(y=colnames(d[1]), x =colnames(d[2]))
  }
  
  else if (!(is.numeric(d[[1]])|is.numeric(d[[2]])))
    
  {
    d %>% ggplot(aes(x = d[[1]], fill = d[[2]]))+
      geom_bar(position = 'dodge')+
      labs(x = names(d)[1], fill = names(d)[2])
  }
  
  else 
    
  {
    print('This function cannot visualize your data.')
  }},
        error=function(e) {
            print('This function cannot visualize your data.')
            print(e)
        }
    )
}

# Test case: A working function should return a box plot with Age along the x-axis and the categories for embarked along the y-axis
visualize(df %>% select(Age, Embarked))

# Test case: A working function should return a box plot with Age along the x-axis and the categories for Age along the y-axis
visualize(df %>% select(Embarked, Age))

# Test case: A working function should return a bar chart of the Sex and Embarked variables
visualize(df %>% select(Sex, Embarked))

# Test case: A working function should return a scatter plot of the Age and Fare variables
visualize(df %>% select(Age, Fare))

# Test Case: I could not think of an example to trigger the 'else' part of the function, given that all columns in the df are either numeric or they are not. To handle this, I decided to incorporate exception handling using tryCatch, where if the user enters an incorrect argument that will raise an error, the function will print 'This function cannot visualize your data.' followed by the error message. Given this, for this test case a working function should return 'This function cannot visualize your data.', followed by an error message
visualize(df %>% select(Age))
```

-------

#### NOTICE: Question 7, 8 and 9 are optional. 

-------

7. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
  
  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the accuracy of the decision model `rpart` where the training data is train_percent. 
```{r}

```

-------

8. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
  
  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the plot of variable important by random forest trained by caret. 

```{r}

```

---
9. Write the following function. Give examples to test your function. 

- Input: 

  - a data frame that has a text column 
  
  - the name of the text column in the data

- Output: the word cloud plot of the text column

Sample codes

- [Word Cloud codes](https://bryantstats.github.io/math421/assignments/text_mining_sample_codes2.html#Total_Word_Cloud)

- [A similar function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#10)
```{r}

```


