
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Handle Missing Values"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.   
```{r}
# Importing data
library(tidyverse)
df <- read_csv("adult_census.csv")

# Showing the number of NAs for each of the columns
colSums(is.na(df))
```

2. Using function `aggr` the VIM package to plot the number of NAs for each column. 
```{r}
# install.packages('VIM')
library(VIM)
aggr(df)
```

3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 
```{r}
# Checking other missing value forms for age variable
table(df$age)
```
```{r}
# Checking other missing value forms for workclass variable
table(df$workclass)

# Can see that '?' represents missing values
```
```{r}
# Checking other missing value forms for education variable
table(df$education)
```
```{r}
# Checking other missing value forms for education.num variable
table(df$education.num)
```
```{r}
# Checking other missing value forms for marital.status variable
table(df$marital.status)
```
```{r}
# Checking other missing value forms for occupation variable
table(df$occupation)

# Can see that '?' represents missing values
```
```{r}
# Checking other missing value forms for relationship variable
table(df$relationship)
```
```{r}
# Checking other missing value forms for race variable
table(df$race)
```
```{r}
# Checking other missing value forms for sex variable
table(df$sex)
```
```{r}
# Checking other missing value forms for capital.gain variable
table(df$capital.gain)
```
```{r}
# Checking other missing value forms for capital.loss variable
table(df$capital.loss)
```
```{r}
# Checking other missing value forms for hours.per.week variable
table(df$hours.per.week)
```
```{r}
# Checking other missing value forms for native.country variable
table(df$native.country)

# Can see that '?' represents missing values
```
```{r}
# Checking other missing value forms for income variable
table(df$income)
```

4. Replace all the forms of missing values found to NA
```{r}
# Only need to account for '?' for other missing value form
df <- replace(df, df == '?', NA)
```

5. Re-plot the number of NAs for each column. 
```{r}
library(VIM)
aggr(df)
```

6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?
```{r}
df_1 <- df %>% drop_na

nrow(df)
nrow(df_1)
```

7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}
df_2 <- df %>% fill(names(df),.direction = 'updown')
colSums(is.na(df_2))
```

8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 
```{r}
colSums(is.na(df))
str(df)
```
```{r}
# Replacing NAs for missing variables (workclass, occupation, native.country)
df_3 <- df
majority_workclass <- names(which.max(table(df_3$workclass)))
df_3$workclass <- replace_na(df_3$workclass, majority_workclass)

majority_occupation <- names(which.max(table(df_3$occupation)))
df_3$occupation <- replace_na(df_3$occupation, majority_occupation)

majority_native.country <- names(which.max(table(df_3$native.country)))
df_3$native.country <- replace_na(df_3$native.country, majority_native.country)

colSums(is.na(df_3))
```
```{r}
# There are no numeric variables with missing values, however, the following...
# ...code would be how to replace NAs with the median

##### median_variable_name <- median(df$variable_name, na.rm=TRUE)
##### df$variable_name <- replace_na(df_variable_name, median_variable_name)
```





