
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 16: Writing a function - Loop"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment16.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------
```{r}
# Copying over functions from previous assignment so that I can knit this assignment
mean_impute <- function(x)
{
mean_of_x <- mean(x, na.rm = TRUE)
library(tidyr)
x <- replace_na(x, mean_of_x)
return(x)
}
```

1. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: a data frame with all the missing of numeric variables replaced by the associated means.

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#28)
```{r}
numeric_df_impute <- function(d){
  for (i in 1:length(d))
  {
      d[[i]] <- mean_impute(d[[i]])
  }
  return(d)
}

# Test Case: Created a numeric data frame with four missing values to be imputed by the function. A working function should impute the four missing values so that the result goes from 1, 1, 2 to 0, 0, 0 when using the colSums function
a = c(1, 2, NA, 3)
b = c(4, NA, 5, 6)
c = c(7, NA, NA, 8)
df1 = data.frame(a,b,c)
colSums(is.na(df1))

df2 <- numeric_df_impute(df1)
colSums(is.na(df2))

```
```{r}
df2
```

-------

2. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: a data frame with all the missing of variables replaced by the associated means (for numeric variables) or modes (for non-numeric variables).

Hint: Use If-statement to combine the function in Problem 1 and the function in [this example](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#28)
```{r}
mode_impute <- function(x)
{
  if(!is.numeric(x))
  {
    # Find the mode of x
    mode_of_x <- names(sort(-table(x)))[1]
    # Replace the missing by the mode
    library(tidyr)
    x <- replace_na(x, mode_of_x) 
  }
return(x)    
}

df_impute <- function(d)
{for (i in 1:length(d))
{if (is.numeric(d[[i]]))
{
  d[[i]] <- mean_impute(d[[i]])
}
 else
 {
   d[[i]] <- mode_impute(d[[i]])
 }
}
 return(d) 
}

# Test Case: Created a data frame with mixed types and five missing values to be imputed by the function. A working function should impute the five missing values so that the result goes from 1, 1, 2, 1 to 0, 0, 0, 0 when using the colSums function
a = c(1, 2, NA, 3)
b = c(4, NA, 5, 6)
c = c(7, NA, NA, 8)
d = c('Red', 'Blue', NA, 'Blue')
df3 = data.frame(a,b,c,d)
colSums(is.na(df3))

df4 <- df_impute(df3)
colSums(is.na(df4))
```
```{r}
df4
```


-------

3. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: Bar plots of all non-numeric variables

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#33)
```{r}
non_num_plots <- function(d)
{library(ggplot2)
  for (i in 1:length(d))
{
  if(!is.numeric(d[[i]]))
  {print(ggplot(data = d, aes(x = d[[i]])) + geom_bar() + labs(x = names(d)[i]))
  }
}
}

# Decided to add more columns and create a new df to make the visuals a bit easier to see and understand when using this function (originally used Titanic df and the visuals were very crowded because of variables such as Name for example)
age = c(9, 13, 16, 19, 22, 26, 28, 29, 32, 35, 43, 56, 67, 78, 83)
sex = c('Male', 'Female', 'Female', 'Male', 'Female', 'Male', 'Male', 'Female', 'Female', 'Female', 'Male', 'Female', 'Male', 'Female', 'Male')
num_grade = c(94, 85, 86, 96, 97, 77, 93, 75, 96, 83, 85, 87, 86, 99, 74)
letter_grade = c('A', 'B', 'B', 'A', 'A', 'C', 'A', 'C', 'A', 'B', 'B', 'B', 'B', 'A', 'C')
absent_ind = c('N', 'Y', 'Y', 'N', 'N', 'Y', 'N', 'Y', 'Y', 'Y', 'Y', 'Y', 'Y', 'N', 'Y')
num_absent = c(0, 1, 2, 0, 0, 4, 0, 5, 1, 2, 1, 1, 3, 0, 6)

df5 = data.frame(age, sex, num_grade, letter_grade, absent_ind, num_absent)

# Test case: A working function should return bar plots of all of the non-numeric variables in the Titanic data set (Name, Sex, Ticket, Cabin, & Embarked)
non_num_plots(df5)
```


-------

4. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: all possible the bar plots of a non-numeric variable filled by a non-numeric variable.

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#40)
```{r}
all_possible_bar_plots <- function(d)
{library(ggplot2)
  for (i in 1:(length(d)-1))
  for (j in (i+1):length(d))
{if ((!is.numeric(d[[i]])) & (!is.numeric(d[[j]])))
{print(ggplot(data = d, aes(x= d[[i]], fill= d[[j]]))+ geom_bar() + 
         labs(x= names(d)[i], fill = names(d)[j]))
 print(ggplot(data = d, aes(x= d[[j]], fill= d[[i]]))+ geom_bar() + 
         labs(x= names(d)[j], fill = names(d)[i]))
}
}
}


# Test case: A working function will produce all of the possible bar plots from the df5 data in which a non-numeric variable can be plotted while having the bars filled with another non-numeric variable (this includes flipped pairs [i.e., a bar plot of sex filled by letter_grade and a bar plot of letter_grade filled by sex])
all_possible_bar_plots(df5)
```

-------

5. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: 

  - all possible the bar plots of a non-numeric variable filled by a non-numeric variable.
  
  - all possible the density plots of a numeric variable colored by a non-numeric variable
  
  - all possible the scatter plots.

Hint: Combine [this function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#37), [this function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#40), and the function in Question 4. One way to combine is creating a new function, `quick_plot`, and call these three functions within `quick_plot`. 
```{r}
quick_plot <- function(d)
{library(ggplot2)
 for (i in 1:(length(d)-1))
  for (j in (i+1):length(d))
{if ((!is.numeric(d[[i]])) & (!is.numeric(d[[j]])))
{print(ggplot(data = d, aes(x= d[[i]], fill= d[[j]]))+ geom_bar() + 
         labs(x= names(d)[i], fill = names(d)[j]))
 print(ggplot(data = d, aes(x= d[[j]], fill= d[[i]]))+ geom_bar() + 
         labs(x= names(d)[j], fill = names(d)[i]))
  
}
  else if ((!is.numeric(d[[i]])) & (is.numeric(d[[j]])))
    {
    print(ggplot(d, aes(x = d[[j]], color = d[[i]]))+ 
              geom_density()+labs(x = names(d)[j], color = names(d)[i]))
    }
  else if ((is.numeric(d[[i]])) & (!is.numeric(d[[j]])))
  {
    print(ggplot(d, aes(x = d[[i]], color = d[[j]]))+ 
              geom_density()+labs(x = names(d)[i], color = names(d)[j]))  
  }
  else if ((is.numeric(d[[i]])) & (is.numeric(d[[j]])))
    {
    print(ggplot(d, aes(x = d[[i]], y = d[[j]]))+ 
              geom_point()+
              labs(x = names(d)[i], y = names(d)[j]))
    print(ggplot(d, aes(x = d[[j]], y = d[[i]]))+ 
              geom_point()+
              labs(x = names(d)[j], y = names(d)[i]))
    }
  }
}

# Test case: A working function will produce all of the possible bar plots in which a non-numeric variable can be filled by another non-numeric variable (expecting 6 of these based on df5 variables), all of the possible density plots in which a numeric variable will be colored by a non-numeric variable (expecting 9 of these based on df5 variables), and all of the possible scatter plots (comprised of 2 numeric variables) (expecting 6 of these based on df5 variables)
quick_plot(df5)
```

