---
title: "Fighter Characteristics"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    logo: UFC_Logo1.png
    source_code: embed
    social: menu

runtime: shiny
---
<style>                     
.navbar {
  background-color:black;
  border-color:black;
}
.navbar-brand {
color white!important;
}
</style>   
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(knitr)
library(DT)

df <- read_csv('ufc_fighter_data.csv')
df$fighter_dob <- as.Date(df$fighter_dob, "%m/%d/%Y")

# Date_range cutoff values may look arbitrary but are just the cutoffs for the ages in days (365 * # of years)
df <- df %>% mutate(total_fights = rowSums(df[ , c(10,11,12)], na.rm=TRUE), 
                    win_pct = fighter_w/total_fights, 
                    over_under_500 = case_when(win_pct >= .500 ~ 'Over', TRUE ~ 'Under'), 
                    weight_range = case_when(fighter_weight_lbs >= 251 ~ '251-300',
                                             fighter_weight_lbs >= 201 ~ '201-250',
                                             fighter_weight_lbs >= 151 ~ '151-200',
                                             fighter_weight_lbs >= 100 ~ '100-150'),
                    age = as.numeric(floor((Sys.Date() - fighter_dob)/365)),
                    age_range = case_when (age >= 51 ~ '51-60',
                                           age >= 41 ~ '41-50',
                                           age >= 31 ~ '31-40',
                                           age >= 19 ~ '19-30')) %>% 
                      select(-fighter_nc_dq, -fighter_nickname, -fighter_id, -fighter_url)

df = drop_na(df)
```

{.sidebar}
=======================================================================
### Ultimate Fighting Championship (UFC)

The Ultimate Fighting Championship (UFC) was founded in November 1993 and holds 
a rich history of Mixed Martial Arts (MMA) combat. This data contains the characteristics of UFC fighters spanning from 1994 to 2023.

```{r}
sliderInput(inputId = "fighter_height_input",
            label = h2("Select Height Range"),
            min = min(df$fighter_height_cm, na.rm = TRUE),
            max = max(df$fighter_height_cm, na.rm = TRUE),
            value= c(152.4, 210.8))
```


Body Measurments
=======================================================================
Row {data-width=500}
-----------------------------------------------------------------------

### Height & Weight

```{r}
renderPlotly({
  p <- df %>% filter(fighter_height_cm>input$fighter_height_input[1], fighter_height_cm<input$fighter_height_input[2]) %>% ggplot(aes(x=fighter_height_cm, y=fighter_weight_lbs, color = over_under_500)) + geom_point() + theme_dark() + ggtitle("Comparison of Fighter Height (cm),\nWeight (lbs), & Record") + labs(caption = "Based on 1994-2023 UFC Fighter Data") + theme(plot.caption = element_text(hjust=1)) + theme(plot.title = element_text(hjust=0.5)) + theme(plot.title = element_text(size=10))
  
ggplotly(p)
})
```

Row {data-width=500}
-----------------------------------------------------------------------

### Reach

```{r}
renderPlotly({
  p2 <- df %>% filter(fighter_height_cm>input$fighter_height_input[1], fighter_height_cm<input$fighter_height_input[2]) %>% ggplot(aes(x=fighter_reach_cm, color = over_under_500)) + geom_density() + theme_dark() + ggtitle("Comparison of Fighter Reach (cm)\nby Record") + labs(caption = "Based on 1994-2023 UFC Fighter Data") + theme(plot.caption = element_text(hjust=1)) + theme(plot.title = element_text(hjust=0.5)) + theme(plot.title = element_text(size=10))
  
ggplotly(p2)
})
```

Fighter Style & Experience
=======================================================================
Column {data-width=500}
-----------------------------------------------------------------------
### Style

```{r}
renderPlotly({
  p3 <- df %>% filter(fighter_height_cm>input$fighter_height_input[1], fighter_height_cm<input$fighter_height_input[2]) %>% ggplot(aes(x=fighter_stance, fill=over_under_500)) + geom_bar() + ggtitle("Number of Fighters by Fighter Stance\n& Record")+ labs(caption = "Based on 1994-2023 UFC Fighter Data") + theme(plot.caption = element_text(hjust=1)) + theme(plot.title = element_text(hjust=0.5)) + theme(plot.title = element_text(size=10)) + ylab('# of Fighters')
  
ggplotly(p3)
})
```

Column {data-width=500}
-----------------------------------------------------------------------
### Experience

```{r}
renderPlotly({
  p4 <- df %>% filter(fighter_height_cm>input$fighter_height_input[1], fighter_height_cm<input$fighter_height_input[2]) %>% ggplot(aes(x=age, y=total_fights, color = over_under_500)) + geom_point() + ggtitle("Comparison of Age,\nTotal Fights, & Record")+ labs(caption = "Based on 1994-2023 UFC Fighter Data") + theme(plot.caption = element_text(hjust=1)) + theme(plot.title = element_text(hjust=0.5)) + theme(plot.title = element_text(size=10))
  
ggplotly(p4)
})
```


Data
=======================================================================
```{r}
datatable(df, options = list(pagelength = 25))
```
